## Statements

```
library(tidyverse)
```
### 1. ```if, else, else if```

```
if (condition) statement
```

```{r}
number <- 1
if (number > 0) "positive number"
```

Use {} for expressions with multiple lines:

```{r}
number <- 1
if (number > 0) {
  "positive number"
}
```

If the condition is not true:

```
if (condition) statement else statement
```

```{r}
number <- -3
if (number > 0) {
  "positive number"
} else {
  "negative number or zero"
}
```

Multiple conditions:

```{r}
number <- 0
if (number > 0) {
  "positive number"
} else if (number < 0){
  "negative number"
} else {
  "zero"
}
```

What if we want to apply the condition to a vector?

```{r}
number <- -2:2
if (number > 0) {
  "positive number"
} else if (number < 0){
  "negative number"
} else {
  "zero"
}
```

Not a success! How can we fix this?

### 2. ```for``` loops

```
for(variable in sequence) statement
```

```{r}
for (i in number) {
  if (i > 0) {
    print("positive number")
  } else if (i < 0) {
    print("negative number")
  } else {
    print("zero")
  }
}
```

For-loops are not commonly used, because there are alternative ways to do the same.

+ Vectorized functions:

```
cumsum(1:10) #cumulative sum
```

+ Functions from the ```apply()``` family that we have seen before.

When you use for-loops, control the length of the object to speed up the computation:

```{r}
number_descriptions <- character(length(number)) #creating a string vector with the same length as the length of the initial vector
for (i in 1:length(number)) {
  if (number[i] > 0) {
    number_descriptions[i] <- "positive number"
  } else if (number[i] < 0) {
    number_descriptions[i] <- "negative number"
  } else {
    number_descriptions[i] <- "zero"
  }
}
number_descriptions
```

### 3. Vectorized conditional statements: ```ifelse()``` and ```dplyr::case_when()```

```
ifelse(number > 0, "positive number", "negative number or zero")
```

More speedy alternatives:

+ dplyr package

```{r}
dplyr::if_else(number > 0, "positive number", "negative number or zero")
```

+ data.table package

```{r}
data.table::fifelse(number > 0, "positive number", "negative number or zero")
```
Multiple conditions:

```{r}
ifelse(number > 0,
       "positive number",
       ifelse(number < 0, "negative number", "zero"))
```

```{r}
dplyr::case_when(
  number > 0 ~ "positive number",
  number < 0 ~ "negative number",
  number == 0 ~ "zero")
```

```{r}
data.table::fcase(
  number > 0, "positive number",
  number < 0, "negative number",
  number == 0, "zero")
```