## Functions 0

```
library(tidyverse)
```
### 1. Creating functions

```{r}
pow <- function(x, p) {
  power <- x ^ p
  return(power)
}
pow(3, 2)
```

If there is no ```return()```, the last computed value is returned:

```{r}
pow <- function(x, p) {
  x ^ p
}
pow(3, 2)
```

What is the problem here?

```{r}
pow <- function(x, p) {
  power <- x ^ p 
}
pow(3, 2) 
```

If the function is short, you don't have to use {}:

```{r}
pow <- function(x, p) x ^ p
pow(3, 2) 
```

Default parameters:

```{r}
pow <- function(x, p = 2) x ^ p
pow(3) 
pow(3, 3)
```

Lazy evaluations:

R is a procrastinator, it reads the the data only when it needs it for computation purposes, this means that there won't be an error if we execute the following piece of code:

```{r}
pow <- function(x, p = 2, we_will_not_use_this_parameter) x ^ p
pow(x = 3)
```

### 2. Adequacy (sanity check) 

We need to make sure that our functions don't produce results for inadequate entry data. For instance, to avoid situations [like that](https://genomebiology.biomedcentral.com/articles/10.1186/s13059-016-1044-7).

What does ```imt()``` do?

```{r}
imt <- function(weight, height) weight / height ^ 2
```

Testing the function:

```{r}
w <- c(60, 80, 120)
h <- c(1.6, 1.7, 1.8)
imt(weight = w, height = h)
```

Do you foresee any problems with the use of this function?

```{r}
imt <- function(weight, height) {
  if (any(height > 3)) warning("The height value is bigger than 3: perhaps, the height value is in cm, not in m (as required)!\n")
  weight / height ^ 2
}
imt(52, 167)
```

What about zero values for weight and height?

```{r}
imt <- function(weight, height) {
  if (any(weight <= 0 | height <= 0)) stop("Body mass index cannot be computed for negative or zero values")
  if (any(height > 3)) warning("The height value is bigger than 3: perhaps, the height value is in cm, not in m (as required)!\n")
  weight / height ^ 2
}
```

[Please look at the debugging techniques](https://adv-r.hadley.nz/debugging.html)

### 3. When and why do we need functions?

+ [Rule of three](https://en.wikipedia.org/wiki/Rule_of_three_(computer_programming))
+ not to repeat the same
+ a way of abstraction

### 4. Functions as objects of the first order

+ Almost eveything in R is a function, even {

+ Functions are objects of the first order, which means that you do with functions many things that you can do with vectors, dataframes etc.:

```{r}
list(mean, min, `{`)
```

+ Recursion: we can use functions are arguments for other functions, we can save functions as variables

+ The same will be true of Python as well (barring the syntactic conventions)

### 5. The family of ```apply()``` functions

+ Working with a matrix

```{r}
A <- matrix(1:12, 3, 4)
A
```

The syntax: ```apply(X, MARGIN, FUN, ...)```, where X — the data, MARGIN is 1 (for rows), 2 (for columns), c(1,2) for rows and columns (= for each element), and FUN — a function that you would like to apply

```{r}
apply(A, 1, sum) #sum for every row #note that a function apply applies to a function sum
apply(A, 2, sum) #sum for every column
apply(A, c(1,2), sum) #sum of every element -- does not make much sense 
```

Additional arguments can be listed:

```{r}
apply(A, 1, sum, na.rm = TRUE)
apply(A, 1, weighted.mean, w = c(0.2, 0.4, 0.3, 0.1)) 
```

+ Anonymous functions

Functions that are used one time and do not have a name

```{r}
apply(A, 1, function(x) sum((x - mean(x))^2)) #sd from the mean by row
apply(A, 2, function(x) sum((x - mean(x))^2)) #sd from the mean by column
apply(A, c(1, 2), function(x) sum((x - mean(x))^2)) #diviation from each value, i.e., zero
apply(A, 1, function(whatevername) sum((whatevername - mean(whatevername))^2)) #you can name the variable in any way you want
```

+ Other functions of the family

  + ```lapply``` -- apply to every element of a list
  
```{r}
some_list <- list(some = 1:10, list = letters)
lapply(some_list, length)
```

  + ```sapply()``` - a version of ```lapply()``` that tries to return the result in a vector or a matrix

```{r}  
sapply(some_list, length)
```

  + applying ```sapply()``` to a vector:
  
```{r}
sapply(1:10, sqrt)
sqrt(1:10)
```

  + we can use ```lapply()``` and ```sapply()``` for dataframes, because a dataframe is basically a list of vectors of the same length, the functions will iterate through columns:
  
```{r}
heroes <- read.csv("https://raw.githubusercontent.com/dashapopova/Intro-to-R/main/week4/heroes_information.csv", 
                   na.strings = c("-", "-99"))
sapply(heroes, class)
```

  + replicate() -- to replicate an operation
  
```{r}
samp <- rlnorm(30)
hist(samp)
```

```{r}
sampdist <- replicate(1000, mean(rlnorm(30)))
hist(sampdist)
```

[A tutorial](https://www.datacamp.com/community/tutorials/r-tutorial-apply-family)


