## Functions

```
library(tidyverse)
```
### 1. Logical operations

**Logical AND**

```{r}
TRUE & TRUE
TRUE & FALSE
FALSE & TRUE
FALSE & FALSE
```

**Logical OR**

```{r}
TRUE | TRUE
TRUE | FALSE
FALSE | TRUE
FALSE | FALSE
```

**Logical NEG**

```{r}
!TRUE
!FALSE
```

**Comparison**

```{r}
2 == 2
2 == 5
2 != 5
2 != c(2, 3)
'flower' == 'rose'
'rose' == 'rose'
c(2,3) == c(2,3)
c(2,5) == c(2,3)
```

### 2. Creating your own functions

```{r}
n_syllables <- function(x){
  stringr::str_count(x, "[АОУЁЕЭЯИЮаоуёеэяию]")
}

n_syllables("корова")

n_syllables("слон")
```

Functions depend on the packages that are loaded, you might want to indicate the necessary package explicitly:

```{r}
n_syllables <- function(x){
  stringr::str_count(x, "[АОУЁЕЭЯИЮаоуёеэяию]")
}
```

Recursion:

```{r}
factorial <- function(x){
  if(x-1 > 0){
    x*factorial(x-1)
  } else {
    1
  }
}

factorial(3)

factorial(4)

factorial(8)

factorial(0)
```

Problem 1:

create a function that will compare the length of two words and tell which one is longer.

To test:

```
which_is_longer("роза", "цветок")
which_is_longer("олень", "палка")
which_is_longer("воробей", "птица")
which_is_longer("boat", "sea")
```

### The use of conditions

Let's take [the data on the UK pubs](https://raw.githubusercontent.com/dashapopova/Intro-to-R/main/assignment1/UK_pubs.csv). Some of the pub names have *Inn* as part of them. Let's plot twenty most common pubs with *Inn* in their names and without *Inn* in their names. 

```{r}
library("dplyr")
library("tidyverse")
```

```{r}
uk_pubs <- read.csv("https://raw.githubusercontent.com/dashapopova/Intro-to-R/main/assignment1/UK_pubs.csv")
uk_pubs %>% 
  count(pub_name, sort = TRUE) %>% 
  mutate(inn = str_detect(pub_name, "Inn")) %>% 
  group_by(inn) %>% 
  slice(1:20) %>% 
  ggplot(aes(fct_reorder(pub_name, n), n))+
  geom_col()+
  coord_flip()+
  facet_wrap(~inn, scale = "free")+
  labs(x = "", y = "", caption = "https://pudding.cool/2019/10/pubs/")
```

We got what we wanted, but the names TRUE and FALSE are not very appealing. We can rectify the situation by using ```ifelse()``` function that has three arguments:

+ condition,
+ the value if the condition takes the value TRUE,
+ the value if the condition takes the value FALSE.

```{r}
ifelse(2+2 == 4, "correct", "incorrect")
ifelse(2+2 == 5, "correct", "incorrect")
```

Let's modify our initial code using this function:

```{r}
uk_pubs %>% 
  count(pub_name, sort = TRUE) %>% 
  mutate(inn = ifelse(str_detect(pub_name, "Inn"), 
                      "with 'inn'", 
                      "without 'inn'")) %>%
  group_by(inn) %>% 
  slice(1:20) %>% 
  ggplot(aes(fct_reorder(pub_name, n), n))+
  geom_col()+
  coord_flip()+
  facet_wrap(~inn, scale = "free")+
  labs(x = "", y = "", caption = "https://pudding.cool/2019/10/pubs/")
```

What if there are more conditions?

```
case_when(
  condition 1 ~ value x,
  condition 2 ~ value y,
  ...
  condition n ~ value z
)
```

Let's look for more type of names: with *Inn*, *Hotel*, *Bar*, *House* and *Tavern*:

```{r}
uk_pubs %>% 
  count(pub_name, sort = TRUE) %>% 
  mutate(place = case_when(
    str_detect(pub_name, "Inn") ~ "inn",
    str_detect(pub_name, "Hotel") ~ "hotel",
    str_detect(pub_name, "Bar") ~ "bar",
    str_detect(pub_name, "House") ~ "house",
    str_detect(pub_name, "Tavern") ~ "tavern")) %>%
  group_by(place) %>% 
  slice(1:10) %>% 
  ggplot(aes(fct_reorder(pub_name, n), n))+
  geom_col()+
  coord_flip()+
  facet_wrap(~place, scale = "free")+
  labs(x = "", y = "", caption = "https://pudding.cool/2019/10/pubs/")
```

What does the NA variant stand for?

```{r}
uk_pubs %>% 
  count(pub_name, sort = TRUE) %>% 
  mutate(place = case_when(
    str_detect(pub_name, "Inn") ~ "inn",
    str_detect(pub_name, "Hotel") ~ "hotel",
    str_detect(pub_name, "Bar") ~ "bar",
    str_detect(pub_name, "House") ~ "house",
    str_detect(pub_name, "Tavern") ~ "tavern",
    TRUE ~ "other")) %>%
  group_by(place) %>% 
  slice(1:10) %>% 
  ggplot(aes(fct_reorder(pub_name, n), n))+
  geom_col()+
  coord_flip()+
  facet_wrap(~place, scale = "free")+
  labs(x = "", y = "", caption = "https://pudding.cool/2019/10/pubs/")
```

What if you want to apply ```summarise()``` or ```mutate()``` only to some columns? For that, there are functions ```summarise_at()``` or ```mutate_at()```. For instance, let's compute the mean for every column of the dataset ```iris``` that start with the word *Sepal*.

```{r}
iris %>% 
  head()
```

```{r}
iris %>% 
  summarise_at(vars(starts_with("Sepal")), mean)
```

```{r}
iris %>% 
  summarise_at(vars(matches("Sepal")), mean) #corresponds to a regular expression
```

```{r}
iris %>% 
  summarise_at(vars(ends_with("Width")), mean)
```

```{r}
diamonds %>% 
  summarise_at(vars(one_of(c("depth", "price", "carat"))), mean)
```

```{r}
diamonds %>% 
  summarise_if(is.numeric, mean)
```

More examples:

```{r}
diamonds

diamonds %>% 
  mutate_at(vars(one_of(c("depth", "price", "carat"))), sqrt)

diamonds %>% 
  mutate_if(is.numeric, sqrt)
```

You can also use your own function:

```{r}
iris %>% 
  summarise_at(vars(matches("Sepal")), function(x){mean(x)*min(x)})
```

Solution to problem 1

```{r}
which_is_longer <- function(x,y){
  if(stringr::str_count(x, "[A-zА-я]")>stringr::str_count(y, "[A-zА-я]")){
    "the first word is longer"
  } else if (stringr::str_count(x, "[A-zА-я]")<stringr::str_count(y, "[A-zА-я]")){
    "the second word is longer"
  } else {
  "the words are of the same length"
  }
}

which_is_longer("роза", "цветок")
which_is_longer("олень", "палка")
which_is_longer("воробей", "птица")
which_is_longer("boat", "sea")
```