## Lists

```
library(tidyverse)
```
### 1. Lists

```{r}
x <- list(m = "a", n = c(2, 4), o = c("blue", "green", "red"))
x
x$m
x[1]
x[[2]]
x[[2]][2]
```

Outside R, lists are usually stored in .json and .xml formats.

### 2. Converting a list into json and vice versa


JSON:

+ {"key":"value"}
+ ["value_1","value_2",...]

+ package ```jsonlite``` (DO NOT execute ```library(jsonlite)```, because it conflicts with ```tidyverse```)

```{r}
got_chars <- jsonlite::read_json("https://raw.githubusercontent.com/dashapopova/Intro-to-R/main/week3/got_chars.json")
got_chars[1]
```

+ to write into a .json:

```{r}
jsonlite::write_json(got_chars, "got_chars.json", auto_unbox = TRUE)
# the auto_unbox argument allows us to preserve the structure that we see in R
```

### 3. Converting .xml into a list and back

+ .xml:  <...> something </...>

```
<bla>Hello, world!</bla>
```
+ there could be attributes, placed in the opening tag:

```
<ingredient amount="3" unit="glass">milk</ingredient>
<ingredient amount="200" unit="gr">flour</ingredient>
<ingredient amount="1" unit="spoon">sugar</ingredient>
```

To read from or to write into an .xml file you can by using ```read_xml()``` and ```write_xml()``` functions from the package ```xml2```:

```{r}
library("xml2")
xml_example <- read_xml('<body>
<ingredient amount="3" unit="glass">milk</ingredient>
<ingredient amount="200" unit="gr">flour</ingredient>
<ingredient amount="1" unit="spoon">sugar</ingredient>
</body>')
xml_example
```

```
write_xml(xml_example, "path/to/the/file.xml")
```

+ To present the data as a list:

```{r}
xml_example <- as_list(xml_example)
xml_example
```

+ Now, the variable is a list:

```{r}
xml_example$body[[2]]
```

+ To get attributes and their values:

```{r}
library(magrittr) # needs to be run every time you start R and want to use %>%
library(dplyr) 
```

```{r}
xml_example$body[[2]] %>% 
  attributes()
```

+ To get a concrete attribute:

```{r}
xml_example$body[[2]] %>% 
  attr("amount")
```

### 4. Viewing lists in R

+ ```View(got_chars)```

+ Another option:

```{r}
str(got_chars[1])
```

```{r}
str(got_chars[1], max.level = 2)
```
```{r}
str(xml_example)
```

### 5. Package ```purrr```

[cheat sheet](https://github.com/rstudio/cheatsheets/raw/master/purrr.pdf)

```{r}
#install.packages("purrr")
library(purrr)
```

+ We can now extract values from a list based on the name:

```{r}
got_chars %>% 
  map("name")
```

+ and based on an index:

```{r}
got_chars %>% 
  map(3)
```

+ to kill the complex structure:

```{r}
got_chars %>% 
  map("name") %>% 
  unlist()
```

+ to create a vector:

```{r}
got_chars %>% 
  map_chr("name")
```

```{r}
got_chars %>% 
  map_int("id")
```

```{r}
got_chars %>% 
  map_lgl("alive")
```

+ we can even create a dataframe:

```{r}
got_chars %>% 
  tibble(name = map_chr(., "name"), 
         id = map_dbl(., "id"),
         alive = map_lgl(., "alive"))
```

+ to get all the attribute values:

```{r}
xml_example$body %>% 
  map(attr_getter("amount"))
```

### 6. Filtering

+ If we have logical expressions in the list, we can use them for filtering:

```{r}
got_chars %>% 
  keep("alive") %>% 
  map_chr("name")
```

+ for the dead:

```{r}
got_chars %>% 
  discard("alive") %>% 
  map_chr("name")
```

+ up to the first FALSE:

```{r}
got_chars %>% 
  head_while("alive") %>% 
  map_chr("name")
```

```{r}
got_chars %>% 
  tail_while("alive") %>% 
  map_chr("name")
```

```{r}
got_chars %>% 
  tibble(name = map_chr(., "name"), 
         alive = map_lgl(., "alive"))
```

### 7. Operations

```{r}
my_l <- list(a = 1:3, b = c("a", "b"), c = list("z", c(8, 9)))
my_l
```

+ to flatten:

```{r}
my_l %>% 
  unlist()
```

```{r}
my_l %>% 
  unlist() %>% 
  unname()
```

```{r}
my_l %>% 
  flatten()
```

+ to add a list:

```{r}
list(c(1, 2), c("a", "b", "c")) %>% 
  append(list(new = c(TRUE, FALSE)))
```

```{r}
list(c(1, 2), c("a", "b", "c")) %>% 
  prepend(list(new = c(TRUE, FALSE)))
```

+ to get unique combinations for two lists:

```{r}
list(c("a", "b"), 1:3) %>% 
  cross() %>% 
  str()
```

```{r}
list("a", "b") %>% 
  cross2(1:3) %>% 
  str()
```