## Strings

### 1. To work with strings in R, use:

+ package ```stringr``` (part of ```tidyverse```)
+ package ```stringi```:

```{r}
#install.packages("stringi")
library(tidyverse)
library(stringi)
```
Function of the package ```stringr``` always start with ```str_```, functions of the package ```stringi``` start with ```stri_```.

[A cheat sheet on stringr](https://github.com/rstudio/cheatsheets/raw/master/strings.pdf)

### 2. How to create a string?

```{r}
"Every farmer who owns a donkey beats it."
```

```{r}
'Every farmer who owns a donkey beats it.'
```

+ an empty string

```{r}
""
```

```{r}
''
```

+ changing the type

```{r}
typeof(4:7)
```
```{r}
as.character(4:7)
```

+ built-in vectors

```{r}
letters
```

```{r}
LETTERS
```

```{r}
month.name
```

+ To create random strings

```{r}
stri_rand_strings(n = 10, length = 5:14)
```

+ To shuffle characters

```{r}
stri_rand_shuffle("Harry Potter and the Goblet of Fire")
```

```{r}
stri_rand_shuffle(month.name)
```
+ To generate random text

```{r}
stri_rand_lipsum(n_paragraphs = 2)
```

### 3. Merging and separating strings

+ To merge strings use the ```str_c``` function with the elements to merge divided by a comma:

```{r}
tibble(upper = rev(LETTERS), smaller = letters) %>% 
  mutate(merge = str_c(upper, smaller))
```

If you want to specify the separator, use the ```sep``` argument:

```{r}
tibble(upper = rev(LETTERS), smaller = letters) %>% 
  mutate(merge = str_c(upper, smaller, sep = "_"))
```

+ To separate:

```{r}
tibble(upper = rev(LETTERS), smaller = letters) %>% 
  mutate(merge = str_c(upper, smaller, sep = "_")) %>% 
  separate(col = merge, into = c("column_1", "column_2"), sep = "_")
```

+ Another way to separate a string is to use the ```str_split```function that returns a list:

```{r}
str_split(month.name, "r")
```

### 4. The number of characters

+ Counting the number of characters

```{r}
tibble(mn = month.name) %>% 
  mutate(n_charactars = str_count(mn))
```

+ Truncating a string:

```{r}
tibble(mn = month.name) %>% 
  mutate(mn_new = str_trunc(mn, 6))
```

```{r}
tibble(mn = month.name) %>% 
  mutate(mn_new = str_trunc(mn, 6, side = "left"))
```

```{r}
tibble(mn = month.name) %>% 
  mutate(mn_new = str_trunc(mn, 6, side = "center"))
```

```{r}
tibble(mn = month.name) %>% 
  mutate(mn_new = str_trunc(mn, 3, ellipsis = ""))
```

+ To make a string longer:

```{r}
tibble(mn = month.name) %>% 
  mutate(mn_new = str_pad(mn, 10))
```

```{r}
tibble(mn = month.name) %>% 
  mutate(mn_new = str_pad(mn, 10,  side = "right"))
```

```{r}
tibble(mn = month.name) %>% 
  mutate(mn_new = str_pad(mn, 10,  pad = "."))
```

### 5. Sorting

```{r}
unsorted_latin <- c("I", "♥", "N", "Y")
sort(unsorted_latin)
```

Another way to accomplish the same:

```{r}
str_sort(unsorted_latin)
```

```{r}
str_sort(unsorted_latin, locale = "lt")
```

```{r}
unsorted_cyrillic <- c("я", "i", "ж")
str_sort(unsorted_cyrillic)
```

```{r}
str_sort(unsorted_cyrillic, locale = "ru_UA")
```

Useful functions: ```stringi::stri_locale_list()```; ```stringi::stri_locale_info``` and ```stringi::stri_locale_set```

The list of locales can be found [here](https://en.wikipedia.org/wiki/List_of_ISO_639-1_codes)

### 6. Substrings

```{r}
tibble(mn = month.name) %>% 
  mutate(has_r = str_detect(mn, "r"))
```

```{r}
tibble(mn = month.name) %>% 
  slice(str_which(month.name, "r"))
```

+ To count the number of times a substring occurs within a string:

```{r}
tibble(mn = month.name) %>% 
  mutate(has_r = str_count(mn, "r"))
```

### 7. To change strings

+ To change the register:

```{r}
latin <- "harrY pOTTer and THE GoBlEt Of fIRE"
cyrillic <- "гаРРи ПоТтеР и кУбок ОГня"
str_to_upper(latin)
```

```{r}
str_to_lower(cyrillic)
```

```{r}
str_to_title(latin)
```

+ To extract a substring

```{r}
tibble(mn = month.name) %>% 
  mutate(mutate = str_sub(mn, start = 1, end = 2))
```

```{r}
tibble(mn = month.name) %>% 
  mutate(mutate = str_extract(mn, "r"))
```

+ To extract all of the occurrences, the function returns a list:

```{r}
str_extract_all(month.name, "r")
```

+ To resplace a substring:

```{r}
tibble(mn = month.name) %>% 
  mutate(mutate = str_replace(mn, "r", "R"))
```

```{r}
tibble(mn = month.name) %>% 
  mutate(mutate = str_replace_all(mn, "r", "R"))
```

+ To delete a substring:

```{r}
tibble(month.name) %>% 
  mutate(mutate = str_remove(month.name, "r"))
```

```{r}
tibble(month.name) %>% 
  mutate(mutate = str_remove_all(month.name, "r"))
```

+ To transliterate:

```{r}
stri_trans_general("stringi", "latin-cyrillic")
```

```{r}
stri_trans_general("творожники", "cyrillic-latin")
```

```{r}
stri_trans_general("stringi", "latin-greek")
```

```{r}
stri_trans_general("stringi", "latin-armenian")
```

### 8. Regular expressions

+ To highlight a pattern:

```{r}
str_view_all("Harry Potter and the Goblet of Fire", "i")
```

```{r}
a <- "4$\\2 + 3$ * 5 = 17$? [Yes|no}^{No|maybe]"
str_view_all(a, "$")
```

```{r}
str_view_all(a, "\\$")
```

```{r}
str_view_all(a, "\\*")
```

```{r}
str_view_all(a, "\\.")
```

```{r}
str_view_all(a, "\\+")
```

```{r}
str_view_all(a, "\\?")
```

```{r}
str_view_all(a, "\\(")
```

```{r}
str_view_all(a, "\\)")
```

```{r}
str_view_all(a, "\\|")
```

```{r}
str_view_all(a, "\\^")
```

```{r}
str_view_all(a, "\\[")
```

```{r}
str_view_all(a, "\\]")
```

```{r}
str_view_all(a, "\\{")
```

```{r}
str_view_all(a, "\\}")
```

```{r}
str_view_all(a, "\\\\")
```

+ Classes of symbols

  + \\d – digits 
  + \\D – non-digits

```{r}
str_view_all("два 15 42. 42 15. 37 08 5. 20 20 20!", "\\d")
```

```{r}
str_view_all("два 15 42. 42 15. 37 08 5. 20 20 20!", "\\D")
```

  + \\s – spaces
  + \\S – not spaces

```{r}
str_view_all("два 15 42. 42 15. 37 08 5. 20 20 20!", "\\s")
```

```{r}
str_view_all("два 15 42. 42 15. 37 08 5. 20 20 20!", "\\S")
```

  + \\w – not spaces and not punctuation marks
  + \\W – spaces and punctuation

```{r}
str_view_all("два 15 42. 42 15. 37 08 5. 20 20 20!", "\\w")
```

```{r}
str_view_all("два 15 42. 42 15. 37 08 5. 20 20 20!", "\\W")
```

  + a group of symbols and its complement

```{r}
str_view_all("Harry Potter and the Goblet of Fire", "[iar]")
```

```{r}
str_view_all("Harry Potter and the Goblet of Fire", "[^iar]")
```

  + built-in groups of symbols

```{r}
str_view_all("два 15 42. 42 15. 37 08 5. 20 20 20!", "[0-9]")
```

```{r}
str_view_all("Гарри Поттер и кубок огня", "[а-я]")
```

```{r}
str_view_all("Гарри Поттер и кубок огня", "[А-Я]")
```

```{r}
str_view_all("Гарри Поттер и кубок огня", "[А-я]")
```

```{r}
str_view_all("Harry Potter and the Goblet of Fire", "[a-z]")
```

```{r}
str_view_all("два 15 42. 42 15. 37 08 5. 20 20 20!", "[^0-9]")
```

  + a|b -- or
  
  + [abc] -- one of

  + [^ab] -- anything but
  
  + [a-c] -- range
  
```{r}
str_view_all("Карл у Клары украл кораллы, а Клара у Карла украла кларнет", "лар|рал|арл")
```

```{r}
str_view_all("Везет Сенька Саньку с Сонькой на санках. Санки скок, Сеньку с ног, Соньку в лоб, все — в сугроб", "[Сс].н")
```

  + the beginning of a string and the end of a string:
  
```{r}
str_view_all("от топота копыт пыль по полю летит.", "^о")
```

```{r}
str_view_all("У ежа — ежата, у ужа — ужата", "жата$")
```

  + ```?regex``` to find more

+ Quantification

  + a? -- zero or one
  + a* -- zero or more
  + a+ -- one or more
  + a{n} -- exactly n
  + a{n,} -- n or more
  + a{n,m} -- between n and m
  
```{r}
str_view_all("хорошее длинношеее животное", "еее?")
```
  
```{r}
str_view_all("хорошее длинношеее животное", "ее*")
```
```{r}
str_view_all("хорошее длинношеее животное", "е+")
```
```{r}
str_view_all("хорошее длинношеее животное", "е{2}")
```
```{r}
str_view_all("хорошее длинношеее животное", "е{1,}")
```
```{r}
str_view_all("хорошее длинношеее животное", "е{2,3}")
```
```{r}
str_view_all("Пушкиновед, Лермонтовед, Лермонтововед", "(ов)+")
```
```{r}
str_view_all("беловатый, розоватый, розововатый", "(ов)+")
```
```{r}
str_view_all("Пушкиновед, Лермонтовед, Лермонтововед", "в.*ед")
```
```{r}
str_view_all("Пушкиновед, Лермонтовед, Лермонтововед", "в.*?ед")
```

  + Look arounds
  
    + a(?=c) -- followed by
    + a(?!c) -- not followed by
    + (?<=b)a -- preceded by
    + (?<!b)a -- not preceded by

```{r}    
str_view_all("Карл у Клары украл кораллы, а Клара у Карла украла кларнет", "а(?=р)")
```

```{r}
str_view_all("Карл у Клары украл кораллы, а Клара у Карла украла кларнет", "а(?=[рл])")
```
```{r}
str_view_all("Карл у Клары украл кораллы, а Клара у Карла украла кларнет", "(?<=р)а")
```
```{r}
str_view_all("Карл у Клары украл кораллы, а Клара у Карла украла кларнет", "(?<=[рл])а")
```
```{r}
str_view_all("Карл у Клары украл кораллы, а Клара у Карла украла кларнет", "р(?!а)")
```
```{r}
str_view_all("Карл у Клары украл кораллы, а Клара у Карла украла кларнет", "(?<!а)р")
```
