### Advanced ```tidyverse```

#### 1. Joining multiple dataframes

+ conjoining two structurally similar dataframes: ```bind_rows(), bind_cols()```

```{r}
library("dplyr")
heroes <- read.csv("https://raw.githubusercontent.com/vydra-v-getrax/intro_to_r/refs/heads/main/week%204/heroes_information.csv", na = c("-","-99"))
```

```{r}
dc <- heroes %>%
  filter(Publisher == "DC Comics") %>%
  group_by(Gender) %>%
  summarise(weight_mean = mean(Weight, na.rm = TRUE))
dc
```

```{r}
marvel <- heroes %>%
  filter(Publisher == "Marvel Comics") %>%
  group_by(Gender) %>%
  summarise(weight_mean = mean(Weight, na.rm = TRUE))
marvel
```

```{r}
other_publishers <- heroes %>%
  filter(!(Publisher %in% c("DC Comics","Marvel Comics"))) %>%
  group_by(Gender) %>%
  summarise(weight_mean = mean(Weight, na.rm = TRUE))
other_publishers
```

Combine by rows (the names of the columns should be the same):

```{r}
bind_rows(dc, marvel)
```

Combine by columns:

```{r}
bind_cols(dc, marvel)
```

Multiple tibbles:

```{r}
bind_rows(dc, marvel, other_publishers)


heroes_list_of_df <- list(DC = dc, 
                          Marvel = marvel, 
                          Other = other_publishers)
bind_rows(heroes_list_of_df)

bind_rows(heroes_list_of_df, .id = "Publisher")
```

+ Relational data: ```*_join()```

```{r}
band_members

band_instruments
```

These two tibbles have a column with the same name, which can be used to join the two structures. Such column is called the key. The key has to identify the observations uniquely.

```{r}
band_members %>%
  left_join(band_instruments)

band_members %>%
  left_join(band_instruments, by = "name")

band_members %>%
  left_join(band_instruments2, by = c("name" = "artist"))

band_members %>%
  right_join(band_instruments)

band_members %>%
  full_join(band_instruments)

band_members %>%
  inner_join(band_instruments)

band_members %>%
  semi_join(band_instruments)

band_members %>%
  anti_join(band_instruments)
```
#### 2. Tidy data: ```tidyr::pivot_longer()```, ```tidyr::pivot_wider()```

Wide format:

```
|student|before Intro to R|after intro to R|
|Kirill|65|63|
|Ania|54|51|
|Anastasia|54|52|
```
Long format:

```
|student|timepoint|weight (kg)|
|Kirill|before Intro to R|65|
|Ania|before Intro to R|54|
|Anastasia|before Intro to R|54|
|Kirill|after Intro to R|63|
|Ania|after Intro to R|51|
|Anastasia|after Intro to R|52|
```


+ ```tidyr::pivot_longer()```

+ ```tidyr::pivot_wider()```

```{r}
new_diet <- tibble(
  student = c("Kirill", "Ania", "Anastasia"),
  before_r_course = c(65, 54, 54),
  after_r_course = c(63, 51, 52)
)
new_diet
```

Let's convert wide format into a longer format:

```{r}
new_diet %>%
  tidyr::pivot_longer(cols = before_r_course:after_r_course,
               names_to = "measurement_time", 
               values_to = "weight_kg")
```

And back:

```{r}
new_diet %>%
  tidyr::pivot_longer(cols = before_r_course:after_r_course,
               names_to = "measurement_time", 
               values_to = "weight_kg") %>%
  tidyr::pivot_wider(names_from = "measurement_time",
              values_from = "weight_kg")
```

#### 3. Transforming multiple columns: ```dplyr::across()```

```{r}
#install.packages("tidyr")
library("tidyr")
```

```{r}
heroes %>%
  group_by(Gender) %>%
  summarise(height = mean(Height, na.rm = TRUE),
            weight = mean(Weight, na.rm = TRUE))
```

```{r}
heroes %>%
  group_by(Gender) %>%
  summarise(across(c(Height,Weight), mean))
```

```{r}
heroes %>%
  group_by(Gender) %>%
  summarise(across(c(Height, Weight), mean, na.rm = TRUE))
```

```{r}
heroes %>%
  drop_na(Height, Weight) %>%
  group_by(Gender) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE))
```

```{r}
heroes %>%
  group_by(Gender) %>%
  summarise(across(where(is.character), 
                   function(x) mean(nchar(x), na.rm = TRUE)))
```

```{r}
heroes %>%
  group_by(Gender) %>%
  summarise(across(where(is.numeric), mean, na.rm = TRUE),
            across(where(is.character), 
                   function(x) mean(nchar(x), na.rm = TRUE)))
```

```{r}
heroes %>%
  group_by(Gender) %>%
  summarise(across(c(Height, Weight), 
                   list(minimum = min,
                        average = mean,
                        maximum = max), 
                   na.rm = TRUE))
```

```{r}
heroes %>%
  group_by(Gender) %>%
  summarise(across(
    c(Height, Weight),
    list(
      min = function(x) min(x, na.rm = TRUE),
      mean = function(x) mean(x, na.rm = TRUE),
      max = function(x) max(x, na.rm = TRUE),
      na_n = function(x, ...) sum(is.na(x)))
    )
    )
```

```{r}
heroes %>%
  mutate(across(where(is.character), as.factor))
```

```{r}
heroes %>%
  count(across(where(function(x) n_distinct(x) <= 6)))
```

4. #### Functional programming: ```purrr```

```{r}
lapply(heroes, class)
```

```{r}
library(purrr)
map(heroes, class)

heroes %>%
  map(class)

heroes %>%
  map_chr(class)

heroes %>%
  map_df(class)

heroes %>%
  map_int(function(x) sum(is.na(x)))

heroes %>%
  map_int(~sum(is.na(.))) #function(arg) is replaced by ~, аnd arg is replaced by .
```

#### 5. List columns and nesting: ```nest()```

```{r}
library(tidyr)
heroes %>%
  nest(!Gender)

heroes %>%
  group_by(Gender) %>%
  nest()

heroes %>%
  group_by(Gender) %>%
  nest() %>%
  mutate(dim = map(data, dim))

heroes %>%
  group_by(Gender) %>%
  nest() %>%
  mutate(dim = map(data, dim)) %>%
  unnest(dim)

heroes %>%
  group_by(Gender) %>%
  nest() %>%
  mutate(dim = map(data, dim)) %>%
  unnest_wider(dim, names_sep = "_") 
```


